---
title: "Variable Selection"
author: "Jakraya (Park) Chatsiriwech - Group 4"
date: "2022-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objectives
The aim of this part of the analysis is to run algorithms to select the important variables into the model and confirm our intuition regarding which variables should explain the price of the ride.

Out intuitive model regress **price on distance, Peak, cab_type, Cab_Name, surge_multiplier and Weather**


# Read data into R

```{r}
data <- read.csv(file = "uber_lyft_cleaned_vF.csv", header = TRUE, 
                 stringsAsFactors = TRUE)
```


# Perform forward selection algorithm

```{r}
library(leaps)
fwd_selection <- regsubsets(price ~ ., data, method = "forward", nvmax = 10)
fwd_summary <- summary(fwd_selection)
fwd_summary$which

# Plot RSS, adjusted R^2, BIC
max_adjr2 <- which.max(fwd_summary$adjr2)
min_bic <- which.min(fwd_summary$bic)

par(mfrow = c(3,1))
plot(fwd_summary$rss, type = "l", xlab = "Number of Variables", ylab = "RSS")
plot(fwd_summary$adjr2, type = "l", xlab = "Number of Variables", ylab = "Adjusted R^2")
points(max_adjr2, fwd_summary$adjr2[max_adjr2], col = "red", pch = 20, cex = 2)
plot(fwd_summary$bic, type = "l", xlab = "Number of Variables", ylab = "BIC")
points(min_bic, fwd_summary$bic[min_bic], col = "red", pch = 20, cex = 2)

# Plot of variables selected
plot(fwd_selection, scale = "adjr2")   # distance, cab_type, Cab_Name, surge_multiplier, Weather (no Rainy), windSpeed
plot(fwd_selection, scale = "bic")  # distance, cab_type, Cab_Name, surge_multiplier
plot(fwd_selection, scale = "Cp") # distance, cab_type, Cab_Name, surge_multiplier

# The chosen model
coef(fwd_selection, 7)  # based on adjusted R^2
coef(fwd_selection, 5)  # based on bic
coef(fwd_selection, 4)  # based on Cp
```

# Perform backward selection algorithm

```{r}
bwd_selection <- regsubsets(price ~ ., data, method = "backward", nvmax = 10)
bwd_summary <- summary(bwd_selection)
bwd_summary$which

# Plot RSS, adjusted R^2, BIC
max_adjr2 <- which.max(bwd_summary$adjr2)
min_bic <- which.min(bwd_summary$bic)

par(mfrow = c(3,1))
plot(bwd_summary$rss, type = "l", xlab = "Number of Variables", ylab = "RSS")
plot(bwd_summary$adjr2, type = "l", xlab = "Number of Variables", ylab = "Adjusted R^2")
points(max_adjr2, bwd_summary$adjr2[max_adjr2], col = "red", pch = 20, cex = 2)
plot(bwd_summary$bic, type = "l", xlab = "Number of Variables", ylab = "BIC")
points(min_bic, bwd_summary$bic[min_bic], col = "red", pch = 20, cex = 2)

# Plot of variables selected
plot(bwd_selection, scale = "adjr2")   # distance, cab_type, Cab_Name, surge_multiplier, Weather (no Rainy), windSpeed
plot(bwd_selection, scale = "bic")  # distance, cab_type, Cab_Name, surge_multiplier
plot(bwd_selection, scale = "Cp") # distance, cab_type, Cab_Name, surge_multiplier

# The chosen model
coef(bwd_selection, 7)  # based on adjusted R^2
coef(bwd_selection, 5)  # based on bic
coef(bwd_selection, 4)  # based on Cp
```

# Perform stepwise selection algorithm using AIC

```{r}
null_mod <- lm(price ~ 1, data = data)
full_mod <- lm(price ~ ., data = data)

# Forward selection
fwd_step_selection <- step(object = null_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "forward", trace = 0)
summary(fwd_step_selection)  # Cab_Name, distance, surge_multiplier, cab_type
plot(fwd_step_selection)

# Backward selection
bwd_step_selection <- step(object = full_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "backward")
summary(bwd_step_selection)  # Cab_Name, distance, surge_multiplier, cab_type

# Both direction
both_step_selection <- step(object = null_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "both")
summary(both_step_selection)  # Cab_Name, distance, surge_multiplier, cab_type
```


# Perform best subset selection
Since the adjusted R^2 suggests the inclusion of `Weather` and `windSpeed`, we will test whether these are worth including in the model or not. We will also check on our intuition with `Peak` as well.

```{r}
best_mod <- regsubsets(price ~ Cab_Name + distance + surge_multiplier + cab_type + Weather + windSpeed + Peak, data = data)

par(mfrow = c(1,2))
plot(best_mod, scale = "adjr2")
plot(best_mod, scale = "bic")

coef(best_mod, 7)
coef(best_mod, 5)
```

# Intermediary conclusion

The base model should include `Cab_Name`, `distance`, `surge_multiplier` and `cab_type` as predictors for `price`. We will run an ANOVA to assess whether `Weather` and `windSpeed` are worth including or not.

```{r}
base_mod <- lm(price ~ Cab_Name + distance + surge_multiplier + cab_type, data = data)
anova(base_mod)
summary(base_mod)   # 73.84% adjusted R^2

weather_mod <- lm(price ~ Cab_Name + distance + surge_multiplier + cab_type + Weather + Weather*surge_multiplier, data = data)  # interaction term between weather & distace / surge_multiplier is significant
anova(weather_mod)
summary(weather_mod)  # same adjusted R^2 even with the interaction term

wind_mod <- lm(price ~ Cab_Name + distance + surge_multiplier + cab_type + windSpeed + windSpeed*distance, data = data)   # even including the interaction term doesn't make windSpeed significant
anova(wind_mod)

combo_mod <- lm(price ~ Cab_Name + distance + surge_multiplier + cab_type + Weather + windSpeed + Weather*windSpeed, data = data)   # the added variables are not significant
anova(combo_mod)
```

# Final conclusion
Discard `windSpeed` and `Peak`. We will keep `Weather` for the next analysis on fitting the model.
