---
title: "Variable Seleection Final"
author: "Jakraya (Park) Chatsiriwech - Group 4"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
From the conclusion in the variable selection stage, the important predictors of price are:
1) Cab_Name
2) distance
3) surge_multiplier
4) cab_type

Although `Weather` by itself is not significant, including an interaction term with either surge_multiplier or distance makes it significant. Backward selection also suggests the inclusion of `windSpeed` but it is insignificant when fitted a model.

From the full (spare) variable selection, a number of algorithm agrees that `cloudCover` gives useful information after fitting distance and surge_multiplier. So, we will also try to explore this further at this stage.


# Read data into R

```{r}
data <- read.csv(file = "uber_lyft_cleaned_vF.csv", header = TRUE, stringsAsFactors = TRUE)

raw_data <- read.csv(file = "rideshare_kaggle.csv", header = TRUE, stringsAsFactors = TRUE)
raw_data <- raw_data[!is.na(raw_data$price), ]

# Extract just the cloudCover column
cloudcover <- raw_data$cloudCover
rm(raw_data)   # Remove raw_data to save space

# combine the cloud cover data with the dataframe
data <- cbind(data, cloudcover)
rm(cloudcover)

# Check NAs
# check for NAs in price
for (i in 1:ncol(data)) {
  if (sum(is.na(data[,i])) >= 1) {
    print(paste(names(data)[i], "contains NAs", sep = " "))
  } else {
    print("all good")
  }
}
```  


# Rerun the variable selection algorithm using AIC

```{r}
null_mod <- lm(price ~ 1, data = data)
full_mod <- lm(price ~ ., data = data)

# Forward selection
fwd_step_selection <- step(object = null_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "forward")
summary(fwd_step_selection)  # Cab_Name, distance, surge_multiplier, cab_type
plot(fwd_step_selection)

# Backward selection
bwd_step_selection <- step(object = full_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "backward")
summary(bwd_step_selection)  # Cab_Name, distance, surge_multiplier, cab_type

# Both direction
both_step_selection <- step(object = null_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "both")
summary(both_step_selection)
```

# Rerun the forward selection algorithm with other error measures

```{r}
library(leaps)
fwd_selection <- regsubsets(price ~ ., data, method = "forward", nvmax = 8)
fwd_summary <- summary(fwd_selection)
fwd_summary$which

# Plot RSS, adjusted R^2, BIC
max_adjr2 <- which.max(fwd_summary$adjr2)
min_bic <- which.min(fwd_summary$bic)

par(mfrow = c(3,1))
plot(fwd_summary$rss, type = "l", xlab = "Number of Variables", ylab = "RSS")
plot(fwd_summary$adjr2, type = "l", xlab = "Number of Variables", ylab = "Adjusted R^2")
points(max_adjr2, fwd_summary$adjr2[max_adjr2], col = "red", pch = 20, cex = 2)
plot(fwd_summary$bic, type = "l", xlab = "Number of Variables", ylab = "BIC")
points(min_bic, fwd_summary$bic[min_bic], col = "red", pch = 20, cex = 2)

# Plot of variables selected
plot(fwd_selection, scale = "adjr2")   # distance, cab_type, Cab_Name, surge_multiplier, Weather (no Rainy), windSpeed
plot(fwd_selection, scale = "bic")  # distance, cab_type, Cab_Name, surge_multiplier
plot(fwd_selection, scale = "Cp") # distance, cab_type, Cab_Name, surge_multiplier

# The chosen model
coef(fwd_selection, 7)  # based on adjusted R^2
coef(fwd_selection, 5)  # based on bic and Cp
```

# Rerun the backward selection algorithm with other error measures

```{r}
bwd_selection <- regsubsets(price ~ ., data, method = "backward", nvmax = 8)
bwd_summary <- summary(bwd_selection)
bwd_summary$which

# Plot RSS, adjusted R^2, BIC
max_adjr2 <- which.max(bwd_summary$adjr2)
min_bic <- which.min(bwd_summary$bic)

par(mfrow = c(3,1))
plot(bwd_summary$rss, type = "l", xlab = "Number of Variables", ylab = "RSS")
plot(bwd_summary$adjr2, type = "l", xlab = "Number of Variables", ylab = "Adjusted R^2")
points(max_adjr2, bwd_summary$adjr2[max_adjr2], col = "red", pch = 20, cex = 2)
plot(bwd_summary$bic, type = "l", xlab = "Number of Variables", ylab = "BIC")
points(min_bic, bwd_summary$bic[min_bic], col = "red", pch = 20, cex = 2)

# Plot of variables selected
plot(bwd_selection, scale = "adjr2")   # distance, cab_type, Cab_Name, surge_multiplier, Weather (no Rainy), windSpeed
plot(bwd_selection, scale = "bic")  # distance, cab_type, Cab_Name, surge_multiplier
plot(bwd_selection, scale = "Cp") # distance, cab_type, Cab_Name, surge_multiplier

# The chosen model
coef(bwd_selection, 7)  # based on adjusted R^2  
coef(bwd_selection, 5)  # based on bic and Cp
```

# Rerun best subset selection

```{r}
best_mod <- regsubsets(price ~ Cab_Name + distance + surge_multiplier + cab_type + + windSpeed + cloudcover + Weather, data = data)

par(mfrow = c(1,2))
plot(best_mod, scale = "adjr2")   # without Weather, all except cloudcover
                                  # with Weather, same result but include Weather (cloudy)
plot(best_mod, scale = "bic")  # without Weather, all except windSpeed and cloudcover -- same result with Weather (no Weather included)

coef(best_mod, 7)
coef(best_mod, 5)
```


# Explore the relationship between Weather, windSpeed and cloudcover

```{r}
library(ggplot2)
ggplot(data, aes(x = windSpeed, fill = Weather)) + geom_histogram()
ggplot(data, aes(x = windSpeed, fill = Weather)) + geom_boxplot()
ggplot(data, aes(x = cloudcover, fill = Weather)) + geom_histogram()
ggplot(data, aes(x = cloudcover, fill = Weather)) + geom_boxplot()

ggplot(data, aes(x = windSpeed, y = price, color = Weather)) + 
  geom_smooth(method = "lm")  # clear increasing trend for 'Clear' category
ggplot(data, aes(x = distance, y = price, color = Weather)) + 
  geom_smooth(method = "lm")
ggplot(data, aes(x = surge_multiplier, y = price, color = Weather)) + 
  geom_smooth(method = "lm")
ggplot(data, aes(x = cloudcover, y = price, color = Weather)) + 
  geom_smooth(method = "lm")  # not very helpful
ggplot(data, aes(x = temperature, y = price, color = Weather)) + 
  geom_smooth(method = "lm")
ggplot(data, aes(x = precipIntensity, y = price, color = Weather)) + 
  geom_smooth(method = "lm")
ggplot(data, aes(x = visibility, y = price, color = Weather)) + 
  geom_smooth(method = "lm")
ggplot(data, aes(x = humidity, y = price, color = Weather)) + 
  geom_smooth(method = "lm")

library(corrplot)
corrplot(cor(data[, c(1,2,5,7,8)]))
```


# Removing redundant variables
Given all models find `Weekday`, `Peak`, `temperature`, `precipIntensity`, `humidity` and `visibility` insignificant, we will remove these from the data set.

```{r}
data <- subset(data, select = -c(Weekday, Peak, temperature, precipIntensity, humidity, visibility))
```


# Fit a model with cloudcover

```{r}
cloud_mod <- lm(price ~ Cab_Name + distance + surge_multiplier + cab_type + cloudcover, data)
summary(cloud_mod)
```

# Relevel the Weather factor and refit a model

```{r}
levels(data$Weather)
data$Weather <- relevel(data$Weather, "Cloudy")  # make Clear the base category, then Cloudy and Rainy
data$Weather <- relevel(data$Weather, "Clear")

mod <- lm(price ~ Cab_Name + distance + surge_multiplier + cab_type + windSpeed, data)
summary(mod)
```

# Final conclusion
`windSpeed` and `cloudcover` are also insignificant, so we will also remove them. 
```{r}
data <- subset(data, select = -c(cloudcover, windSpeed))

write.csv(data, file = "uber_lyft_cleaned_vF.csv", row.names = FALSE, col.names = TRUE)
```

