---
title: "Variable Selection (Spare)"
author: "Jakraya (Park) Chatsiriwech - Group 4"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(usethis)
# usethis::edit_r_environ()
```

# Read the raw data into R

```{r}
raw_data <- read.csv(file = "rideshare_kaggle.csv", header = TRUE, stringsAsFactors = TRUE)
raw_data <- raw_data[!is.na(raw_data$price), ]
weather_data <- raw_data[, c(13:15, 18:19, 22:36, 38:57)]  # first 18 columns already in the data_vF

rm(raw_data)

cor_matrix <- cor(weather_data)
# cor_matrix[which(cor_matrix >= 0.9)]
library(corrplot)
corrplot(cor_matrix, type = "lower", tl.cex = 0.5) 

library(dplyr)
weather_data <- subset(weather_data, select = -c(windGustTime, temperatureHighTime, temperatureLowTime, apparentTemperatureHighTime, apparentTemperatureLowTime, sunsetTime, sunriseTime, moonPhase, uvIndexTime, temperatureMinTime, temperatureMaxTime, apparentTemperatureMinTime, apparentTemperatureMaxTime))

cor_matrix <- cor(weather_data)
corrplot(cor_matrix, type = "lower", tl.cex = 0.5) 

weather_data <- subset(weather_data, select = -c(visibility.1, temperatureMax, temperatureMin, apparentTemperatureMin, apparentTemperatureMax))

cor_matrix <- cor(weather_data)
corrplot(cor_matrix, type = "lower", tl.cex = 0.5) 

weather_data <- subset(weather_data, select = -c(apparentTemperature, apparentTemperatureHigh, apparentTemperatureLow, dewPoint))

cor_matrix <- cor(weather_data)
corrplot(cor_matrix, type = "lower", tl.cex = 0.5) 
```

# Forward selection

```{r}
library(leaps)
fwd_selection <- regsubsets(price ~ ., weather_data, method = "forward", nvmax = 10)
fwd_summary <- summary(fwd_selection)

fwd_summary$which

# Plot RSS, adjusted R^2, BIC
max_adjr2 <- which.max(fwd_summary$adjr2)
min_bic <- which.min(fwd_summary$bic)

par(mfrow = c(3,1))
plot(fwd_summary$rss, type = "l", xlab = "Number of Variables", ylab = "RSS")
plot(fwd_summary$adjr2, type = "l", xlab = "Number of Variables", ylab = "Adjusted R^2")
points(max_adjr2, fwd_summary$adjr2[max_adjr2], col = "red", pch = 20, cex = 2)
plot(fwd_summary$bic, type = "l", xlab = "Number of Variables", ylab = "BIC")
points(min_bic, fwd_summary$bic[min_bic], col = "red", pch = 20, cex = 2)

# Plot of variables selected
plot(fwd_selection, scale = "adjr2")   # distance, surge_multiplier, cloudCover
plot(fwd_selection, scale = "bic")  # distance, surge_multiplier
plot(fwd_selection, scale = "Cp") # distance, surge_multiplier, cloudCover

# The chosen model
coef(fwd_selection, 3)  # based on adjusted R^2 and Cp
coef(fwd_selection, 2)  # based on bic
```

# Perform backward selection algorithm

```{r}
bwd_selection <- regsubsets(price ~ ., weather_data, method = "backward", nvmax = 10)
bwd_summary <- summary(bwd_selection)
bwd_summary$which

# Plot RSS, adjusted R^2, BIC
max_adjr2 <- which.max(bwd_summary$adjr2)
min_bic <- which.min(bwd_summary$bic)

par(mfrow = c(3,1))
plot(bwd_summary$rss, type = "l", xlab = "Number of Variables", ylab = "RSS")
plot(bwd_summary$adjr2, type = "l", xlab = "Number of Variables", ylab = "Adjusted R^2")
points(max_adjr2, bwd_summary$adjr2[max_adjr2], col = "red", pch = 20, cex = 2)
plot(bwd_summary$bic, type = "l", xlab = "Number of Variables", ylab = "BIC")
points(min_bic, bwd_summary$bic[min_bic], col = "red", pch = 20, cex = 2)

# Plot of variables selected
plot(bwd_selection, scale = "adjr2")   # distance, surge_multiplier, windSpeed, windGust, temperatureHigh, cloudCover
plot(bwd_selection, scale = "bic")  # distance, surge_multiplier
plot(bwd_selection, scale = "Cp") # distance, surge_multiplier, cloudCover

# The chosen model
coef(bwd_selection, 6)  # based on adjusted R^2
coef(bwd_selection, 2)  # based on bic
coef(bwd_selection, 3)  # based on Cp
```

# Perform stepwise selection algorithm using AIC

```{r}
null_mod <- lm(price ~ 1, data = weather_data)
full_mod <- lm(price ~ ., data = weather_data)

# Forward selection
fwd_step_selection <- step(object = null_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "forward", trace = 0)
summary(fwd_step_selection)  # distance, surge_multiplier, cloudCover (insignificant)
# plot(fwd_step_selection)

# Backward selection
bwd_step_selection <- step(object = full_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "backward")
summary(bwd_step_selection)  # distance, surge_multiplier, windSpeed, windGust, cloudCover

# Both direction
both_step_selection <- step(object = null_mod, 
                           scope = list(lower = null_mod, upper = full_mod),
                           direction = "both")
summary(both_step_selection)  # distance, surge_multiplier, cloudCover
```


# Perform best subset selection
Since the adjusted R^2 suggests the inclusion of `windGust` and `windSpeed`, we will test whether these are worth including in the model or not. We will also check on our intuition with `temperatureHigh` as well.

```{r}
best_mod <- regsubsets(price ~ distance + surge_multiplier + windSpeed + windGust + cloudCover + temperatureHigh, data = weather_data)

par(mfrow = c(1,2))
plot(best_mod, scale = "adjr2")   # include all
plot(best_mod, scale = "bic")  # distance, surge_multiplier

coef(best_mod, 6)
coef(best_mod, 2)
```
